Exploring Zstandard user-provided dictionary compression for FASTA files

### 2022-12-10 paper/paper.qmd
Final code

The model contains two pipelines, one for Zstd custom dictionary generation and another for generating transcoded compression data using Bzip2, Xz, zLib, and Zstd compressors, with a general overview in @fig-model. Julia was chosen as there is package support for working with sequence data as well as for trivial transcoding of compressed or decompressed data via TranscodingStreams.jl wherein compression algorithm implementations are loaded as codecs[@sato]. A list of all direct dependencies can be found in the repository's Project.toml file. The Julia implementation operates by first unloading a tarball containing the set of FASTA training files into a Julia artifact, or life-cycled datastore. The files are plugged into Zstd's training mode, via the Zstd command line interface --train option, to generate a custom dictionary specific to that dataset. Subsequent dictionary manipulation in Julia is based on an experimental CodecZstd.jl branch that includes functions that wrap the Zstd Dictionary API[@kittisopikul]. Random FASTA data is generated using BioSequences.jl[@biosequences] and FASTX.jl[@fastx] APIs which is then transcoded into compressed formats via the CodecBzip2.jl[@codecbzip2], CodecXz.jl[@codecxz], CodecZlib.jl[@codeczlib], and CodecZstd.jl[@codeczstd] codecs for comparative analysis. Both dictionary-based and default Zstd compression are applied to determine if the former desmontrates a significant reduction in data size.

![Overview of the Julia implementation for the compression model. **A**, Custom dictionary generation via an Julia artifact of FASTA files plugged into the Zstd Dictionary API. **B**, Transcoding randomly generated FASTA formatted data using TranscodingStreams.jl[@sato] compression codecs. ](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/figures/Compression_Pipeline_Model.png){#fig-model}

@@

@@

# Supporting information

@@

::: {.supp}
## S1 File. {#s1-file}

:::

::: {.supp}

## S1 Video. {#s1-video}

::: {.supp}
## S1 Appendix. {#s1-appendix}

::: {.supp}
## S1 Table. {#s1-table}

>>>>

The model contains two pipelines, one for Zstd custom dictionary generation and another for generating transcoded compression data using Bzip2, Xz, zLib, and Zstd compressors, with a general overview in @fig-model. Julia was chosen as there is package support for working with sequence data as well as for trivial transcoding of compressed or decompressed data via TranscodingStreams.jl wherein compression algorithm implementations are loaded as codecs[@sato]. A list of all direct dependencies can be found in the repository's Project.toml file. The Julia implementation operates by first unloading a tarball containing the set of FASTA training files into a Julia artifact, or life-cycled datastore. The files are plugged into Zstd's training mode, via the Zstd command line interface --train option, to generate a custom dictionary specific to that dataset that is then plugged again into the CLI for data compression. Random FASTA data is generated using BioSequences.jl[@biosequences] and FASTX.jl[@fastx] APIs which is then transcoded into compressed formats via the CodecBzip2.jl[@codecbzip2], CodecXz.jl[@codecxz], CodecZlib.jl[@codeczlib], and CodecZstd.jl[@codeczstd] codecs for comparative analysis. Both dictionary-based and default Zstd compression are applied to determine if the former results in a significant reduction of overall data size.

![Overview of the Julia implementation for the compression model. **A**, Custom dictionary generation via an Julia artifact of FASTA files plugged into the Zstd Dictionary API. **B**, Transcoding randomly generated FASTA formatted data using TranscodingStreams.jl[@sato] compression codecs.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/diagrams/Compression_Model_Overview.png){#fig-model}

@@

![Average FASTA formatted data size (bytes) following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression. *Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation. Randomized DNA sequences within the range of 100 000 to 150 000 base pairs.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Compression_Performance_Boxplot.png){#fig-performance}

@@

# Supplementary information

@@

### 2022-12-10 paper/paper.tex
Final code

The model contains two pipelines, one for Zstd custom dictionary
generation and another for generating transcoded compression data using
Bzip2, Xz, zLib, and Zstd compressors, with a general overview in
Fig~\ref{fig-model}. Julia was chosen as there is package support for
working with sequence data as well as for trivial transcoding of
compressed or decompressed data via TranscodingStreams.jl wherein
compression algorithm implementations are loaded as codecs\citep{sato}.
A list of all direct dependencies can be found in the repository's
Project.toml file. The Julia implementation operates by first unloading
a tarball containing the set of FASTA training files into a Julia
artifact, or life-cycled datastore. The files are plugged into Zstd's
training mode, via the Zstd command line interface --train option, to
generate a custom dictionary specific to that dataset. Subsequent
dictionary manipulation in Julia is based on an experimental
CodecZstd.jl branch that includes functions that wrap the Zstd
Dictionary API\citep{kittisopikul}. Random FASTA data is generated using
BioSequences.jl\citep{biosequences} and FASTX.jl\citep{fastx} APIs which
is then transcoded into compressed formats via the
CodecBzip2.jl\citep{codecbzip2}, CodecXz.jl\citep{codecxz},
CodecZlib.jl\citep{codeczlib}, and CodecZstd.jl\citep{codeczstd} codecs
for comparative analysis. Both dictionary-based and default Zstd
compression are applied to determine if the former desmontrates a
significant reduction in data size.

@@

{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/figures/Compression_Pipeline_Model.png}

@@

@@

\hypertarget{supporting-information}{%
\section{Supporting information}\label{supporting-information}}

@@

\paragraph*{S1 File.}
\label{s1-file}
{\textbf{Lorem ipsum.}}

\paragraph*{S1 Video.}
\label{s1-video}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{s1-appendix}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\paragraph*{S1 Table.}
\label{s1-table}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

>>>>

The model contains two pipelines, one for Zstd custom dictionary
generation and another for generating transcoded compression data using
Bzip2, Xz, zLib, and Zstd compressors, with a general overview in
Fig~\ref{fig-model}. Julia was chosen as there is package support for
working with sequence data as well as for trivial transcoding of
compressed or decompressed data via TranscodingStreams.jl wherein
compression algorithm implementations are loaded as codecs\citep{sato}.
A list of all direct dependencies can be found in the repository's
Project.toml file. The Julia implementation operates by first unloading
a tarball containing the set of FASTA training files into a Julia
artifact, or life-cycled datastore. The files are plugged into Zstd's
training mode, via the Zstd command line interface --train option, to
generate a custom dictionary specific to that dataset that is then
plugged again into the CLI for data compression. Random FASTA data is
generated using BioSequences.jl\citep{biosequences} and
FASTX.jl\citep{fastx} APIs which is then transcoded into compressed
formats via the CodecBzip2.jl\citep{codecbzip2},
CodecXz.jl\citep{codecxz}, CodecZlib.jl\citep{codeczlib}, and
CodecZstd.jl\citep{codeczstd} codecs for comparative analysis. Both
dictionary-based and default Zstd compression are applied to determine
if the former results in a significant reduction of overall data size.

@@

{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/diagrams/Compression_Model_Overview.png}

@@

{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Compression_Performance_Boxplot.png}

\caption{\label{fig-performance}Average FASTA formatted data size
(bytes) following Bzip2, Xz, Zlib, default Zstd and dictionary-based
Zstd compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
artifact for Zstd dictionary generation. Randomized DNA sequences within
the range of 100 000 to 150 000 base pairs.}

@@

\hypertarget{supplementary-information}{%
\section{Supplementary information}\label{supplementary-information}}

@@
